---
title: "Extraction of demographic metrics"
author: "Ruth Kelly"
date: "20/02/2020"
output: html_document
---

## 

This document is an overarching file which works with the r scripts in the same folder to calculate the demographic metrics used in Kelly et al. 2020, 'Environmental and evolutionary context are required to infer plant life-history strategies from functional traits.'

Required code and datafiles for this code are as follows:

Datafiles
* COMPADRE_v.4.0.1.RData
* Brosimum_alicastrum_27_10_2017_female_matA.csv
* Brosimum_alicastrum_27_10_2017_female_matF.csv
* silence_matrices_ellis_2012.csv
* all_trait_means10_10_2017_added_datasets_vr4.csv
* all_trait_means20_04_2018_added_datasets.csv

R scripts
* checking_matrices_Try_matches_03_2020_clean.R
* merging_matrices_13_03_2020.r
* seasonal_matrices_13_03_2018.R
* long_periods_analysis_13_03_2030.R
* splitting_seed_stage_number_13_03_2020.r
* metrics_annual_no_seeds_13_03_2020.R
* metrics_annual_1_seedstage_13_03_2020.R
* metrics_annual_2_seedstage_13_03_2020.R
* metrics_annual_3_seedstage_13_03_2020.R
* refining_metrics_dataset_13_03_2020.R
* match_demography_and_traits_03_04_2020.R

R functions
* exceptionalLife_function_KHealy_RK_2018.R
* makelifetable_mx0_function_18_04_2018.R
* pSeedling_function_15_06_2017.R

This first chunk of code sets my working directory, change this to wherever these files are stored on your machine. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir = "C:/R/Kelly_et_al_traits_to_demog_2020/1_Calculating_Life_History_Metrics")
```


The following libraries are needed.  If you don't have them, install them first using the function 'install.packages()'

```{r libraries}
library("popdemo")
library("vegan")
library("popbio")
library("dplyr")
library("expm")
library("pracma")
library("MASS")
library("ineq")
library("dplyr")
library("geosphere")

```

### Step 1 Select suitable species/matrices 

This code reads in Compadre_v.4.0.1.RData and selects usuable matrices on the following criteria:
 * Unmanipulated conditions
 * Vascular plant species
 * Has separate A, U and F matrices
 * Sexual fecundity has been measured
 * Clonal fecundity is not included (i.e. new individuals are treated as such even if they are clones, this is for data compatibility purposes as most studies do not diffferentiate between clones and non-clones)
 * Annual transition matrices
 * Sum of surivival in columns of the transition matrix (U) <= 1
 

Furthermore, individual matrices are chosen by preference, where no individual matrices exist for a species in Compadre 'mean' matrices are used (these may have been averaged accross time or space, but not different treatments), finally if no individual or mean matrices exist pooled matrices are used (these may averaged accross different treatments etc. and were supplied as such to Compadre in the primary literature).

The resulting species were then matched to the list of species for which trait data was available for at least 4 traits of interest (in January 2018). This resulted in 180 matching species datasets with sufficient trait data. Each of these studies was then individually error checked to make sure that the Compadre data matched the original paper.  Where necessary ammendments where made to matrices and metadata to match the original sources. 
 
Output of this code is a new RData file in the same format as compadre containing only these selected matrices and entitled :"ammended_annual_compadre_data_13_03_2020.RData"

####

```{r Step 1 Select suitable matrices, include = FALSE}

source("checking_matrices_Try_matches_03_2020_clean.R")
```

### Step 2. Average individual matrices from same study and location

Here we average matrices which were measured on the same populations in multiple years.
This gives a temporally averaged matrix and helps to avoid later issues in the metrics e.g. 'infinite lifespan' or 'never reproducing' which arise as a result of individual 
years where no individuals in a study population died or reproduced respectively. 

Resulting matrices are the checked to make sure they are 
 * Ergodic
 * Irreducible
 * Primitive
 
Matrices which do not meet this criteria are discarded from the dataset. 
 
This code requires the dataset "ammended_annual_compadre_data_13_03_2020.RData" created in the previous step.

This code outputs the dataset "selected_annual_matrices_merged_per_population_13_03_2018.RData" required in the next step.


```{r Temporal averaging of individual matrices}

source("merging_matrices_13_03_2020.r")

```

### Step 3. Calculate metrics for non-annual species

Code to calculate demographic metrics for matrices marked with a periodicity different from yearly in Compadre dataset.  This includes species for which original matrices are supplied are seasonal transitions or multiyear transitions, but also species where the transition matrices are annual according to the original source and incorrectly classified in this Compadre version (because the Annual Periodicity refers incorrectly to the field work schedule and not the transitions in the final matrix). All matrices contained in this section have been checked against the original source and are adjusted as necessary. 

These matrices are then appended to the merged matrices from step 2 ("selected_annual_matrices_merged_per_population_08_06_2018.RData") and are outputted first as "selected_matrices_with_seasonal_merged_per_population_13_03_2020.RData" after seasonal matrices are added and then as "all_selected_matrices_merged_per_population_13_03_2020.RData" after long period matrices are added. 

These take a little while to run, be patient :)

```{r Step 3 Checking matrix adjustments and metric calculation for seasonal matrices}
source("seasonal_matrices_13_03_2020.R")
source("long_periods_analysis_13_03_2020.R")

```



### Step 4. Split matrices by number of seed stages.

Splits annual matrices by the number seed stages, checks that these seed stages are at the beginning and the life-stages in the matrix, and removes matrices with fewer than 2 non-seed stages. 

This code requires the dataset "all_selected_matrices_merged_per_population_03_2020.RData" created in step 2. 

And outputs datasets separated by the number of seed stages in the matrices as : 
* No seedbank recorded = "annual_no_Seedbank_03_2020.RData"
* 1 seedbank stage recorded = "annual_1yr_Seedbank_03_2020.RData"
* 2 seedbank stage recorded = "annual_2yr_Seedbank_03_2020.RData"
* 3 seedbank stage recorded = "annual_3yr_Seedbank_03_2020.RData"


```{r Step 3 Split matrices by number of seed stages}

source("splitting_seed_stage_number_13_03_2020.r")

```

### Step 4.  Calculate life-history metrics for annual transition matrices... 

Seperately, calculates for matrices with different numbers of seed stages, using correction for calculation of annual fecundity, progressive growth and retrogressive growth so that these are calculated from the non-seed stage information only.  This makes studies with unmeasured seed-banks compatible with those where seeds were counted. 

The following metrics are derived for each matrix
* Generation Time
* Mean Progressive Growth (SSD weighted)
* Mean Retrogressive Growth (SSD weighted)
* Mean Reproduction
* Net reproductive rate (SSD weighted)
* Mean age at reproduction
* Mature life expectancy
* Mean life expectancy
* Lifespan (Age 99% dead)
* Survival as mean life/life span
* Gini Index for degree of iteroparity


File outputs = 
* "demography_non_seedbank_03_2020.csv"
* "demography_1yr_seedbank_03_2020.csv"
* "demography_2yr_seedbank_03_2020.csv"
* "demography_3yr_seedbank_03_2020.csv"

Be patient here too! 
 
Some messages regarding errors will appear in the console with you run this chunk. They produce NAs in the dataset from some loops, but these are then fixed or removed in the following steps. It is therefore okay to ignore these messages in the console here. 

```{r Step 3 calculate life-history metrics, echo = FALSE, include = FALSE}

source("metrics_annual_no_seeds_13_03_2020.R")
source("metrics_annual_1_seedstage_13_03_2020.R")
source("metrics_annual_2_seedstage_13_03_2020.R")
source("metrics_annual_3_seedstage_13_03_2020.R")

```

### Step 5.  Checking for implausible outliers based on biology.  

Check demographic metrics for implausible outliers based on species biology, 
remove rows where life_span or Gini index or survival ratio could not be calculated due to survival issues (i.e. no death observed in final life stage in study).

Then calculate mean values per species per GPS location. 

code then looks for more matrices that could be spatially averaged, where species matrices are from the same habitat and within 5km of each other and 
averages these. 

Following this the original code select a random population per species where multiple locations still existed. This has been editted so that the code 
now always selects the same populations, these being those that were originally 
selected randomly and used in the paper. This effects 13 (out of 90) species for which alternative spatial locations could have been chosen. 

Exports all of the important demography information to 

Output file : "demo_means_selected_03_04_2020.csv"


```{r Step 5 combining metrics}
source("refining_metrics_dataset_13_03_2020.R" )


```

### Step 6. Merge demographic metrics with trait dataset. 


Selected species with 4 matching traits, LMA, LA, Height and seed mass. 

Required files: 
"all_trait_means20_04_2018_added_datasets.csv"
"demo_means_selected_03_04_2020.csv"

Output file: 
 -"matching_4traits_demo_data_03_04_2020.csv" 
 
This file contains matching Height, Leaf area, LMA, and Seed mass for all 
species where we had data on all 4 of these traits (80 species)
 
This file feeds into the next steps given in the folder 2_Joining_Env_phylo_and_GLMMing - if you've survived this far, go there next,
follow the overview rmd there :)


```{r Step 6 Merging demographic metrics and trait data}
source("match_demography_and_traits_03_04_2020.R")


```


